---
title: "R Notebook"
output: html_notebook
---

# *Lab component* (90 points total) 
Please refer to module 2 and lab 3 and complete tasks 3---5. 

Suppose public health officials in a small city need to decide how much resources to devote toward prevention and treatment of a certain disease, but the fraction $\theta$ of infected individuals in the city is unknown.

Suppose they allocate enough resources to accomodate a fraction $c$ of the population. If $c$ is too large, there will be wasted resources, while if it is too small, preventable cases may occur and some individuals may go untreated. After deliberation, they tentatively adopt the following loss function:

$$
\ell(\theta, c) = \begin{cases}
              | \theta - c |, & \text{if } c \geq \theta \\
    10 \times | \theta - c |, & \text{if } c < \theta    \\
\end{cases}
$$

By considering data from other similar cities, they determine a prior $p(\theta)$. For simplicity, suppose $\theta\sim\text{Beta}(a,b)$ (i.e., $p(\theta) = \text{Beta}(\theta|a,b)$), with $a=0.05$ and $b=1$. They conduct a survey assessing the disease status of $n=30$ individuals, $x_1,\ldots,x_n$. This is modeled as $X_1,\ldots,X_n \stackrel{iid}{\sim} \text{Bernoulli}(\theta)$, which is reasonable if the individuals are uniformly sampled and the population is large. Suppose all but one are disease-free, i.e., $\sum_{i=1}^n x_i = 1$.

-----

\color{black}

## Task 1: Find minimum cost ($c$) given $a$, $b$, $n$, and $\sum x_i$

We know $p(\theta|x)$ as an updated Beta, so we can numerically compute this integral for each $c$.  Reproduce Figure 1 from lecture, illustrating $\rho(c,x)$ for our example. Also, work through where the minimum occurs numerically ($c\approx 0.08$).

\color{MyColor}

Consider the loss function.
$$
\ell(\theta, c) = \begin{cases}
              | \theta - c |, & \text{if } c \geq \theta \\
    10 \times | \theta - c |, & \text{if } c < \theta    \\
\end{cases}
$$

```{r,cache=TRUE}
# compute the loss given theta and c 
loss_fun = function(theta, c){
    if (c < theta){
        return(10 * abs(theta - c))
    } else{
        return(abs(theta - c))
    } # end ifelse
} # end func
```

We now write a function **posterior_risk** which is a function of cost (c), parameters a_prior and b_prior for the prior distribution of $\theta$, the summation of $x_i$ sum_x, the number of observations n, and also the number of random draws s. 
```{r}
# compute the posterior risk given c (cost) 
# s is the number of random draws 
posterior_risk = function(cost, shape, sum_x, n, s = 30000){
    ### get posterior
    a_prior = shape[1]; b_prior = shape[2]
    a_post  = a_prior + sum_x
    b_post  = b_prior + n - sum_x
    
    ### randow draws from beta distribution 
    theta = rbeta(s, a_post, b_post)
    loss  = apply(as.matrix(theta), 1, loss_fun, cost)
    
    ### average values from the loss function
    risk = mean(loss)
} # end fun 
```

create function to search for the minimum cost
$$\text{cost}^* = \underset{c}{\operatorname{argmin}} \rho(c,x)$$
```{r}
cost_min_risk = function(cost, risk){
    ### check dimension
    stopifnot(length(cost) == length(risk))
    ### return cost created minimum risk
    idx = which.min(risk)
    res = c(cost[idx], risk[idx])
    return(res)
} # end fun
```

calculate posterior risk and reproduce figure 1 in lecture 2
```{r, fig.height=4, fig.width=6}
### data
sum_x = 1
n = 30

### prior parameters
a = 0.05; b = 1

### generate posterior risk
cost = seq(0, 0.5, by = 0.01)
s    = 2000
set.seed(123)
post_risk = apply(as.matrix(cost), 1, posterior_risk, c(a, b), sum_x, n, s)

### find the minimum cost
tmp = cost_min_risk(cost, post_risk)
cost_opt = tmp[1]; risk_min = tmp[2]

### reproduce figure 1
plot(  cost, post_risk,    col = "blue", type = "l", xlab = "Cost", ylab = "Posterior Risk (cost, X)")
points(cost_opt, risk_min, col = "red",  pch = 19)
abline(v = cost_opt, col = "red", lty = 2)
abline(h = risk_min, col = "red", lty = 2)
text(0.4, 0.2,  paste("optimal cost =", cost_opt, "\n", 
                      "minimum risk =", round(risk_min, 3)))
```

-----

\color{black}

## Task 2: Sensitivity Analysis
**Now perform a sensitivity analysis for the prior assumption (Beta(a,b)). What do you find?**

\color{MyColor}

```{r}
### initialization
cost = seq(0, 0.5, by = 0.01)
s    = 2000

### set prior
as = c(0.05, 1, 0.05); bs = c(1, 1, 10)
shape_prior = cbind(as, bs)

### for each pair of a and b, compute the posterior risks
post_risk = apply(shape_prior, 1, function(shape){
    apply(as.matrix(cost), 1, posterior_risk, shape, sum_x, n, s)
}) # end apply
post_risk = t(post_risk)

### plot the results
plot(cost, post_risk[1,], type = 'l', col='blue', lty = 1, yaxt = "n", ylab = "p(c, x)")
par(new = T)
plot(cost, post_risk[2,], type = 'l', col='red', lty = 2, yaxt = "n", ylab = "")
par(new = T)
plot(cost, post_risk[3,], type = 'l', lty = 3, yaxt = "n", ylab = "")

legend("bottomright", lty = c(1,2,3), col = c("blue", "red", "black"), 
       legend = c("a = 0.05 b = 1", "a = 1 b = 1", "a = 0.05 b = 5"))
```

-----

\color{black}

## Task 3
**Consider the Bayes procedure ($c\approx 0.08$), $c=\bar{x}, c=0.1.$ Reproduce Figure 2. Explain your findings. **

\color{MyColor}


```{r}
### data
sum_x = seq(0, 30, by = 1)
n = 30

### prior parameters
a = 0.05; b = 1

### generate posterior risk
cost = seq(0, 1.0, by = 0.01)
s    = 2000
set.seed(123)
cost_opts = sapply(sum_x, function(x){
    ### generate posterior risk
    post_risk = apply(as.matrix(cost), 1, posterior_risk, c(a, b), x, n, s)
    
    ### get optimal cost
    tmp = cost_min_risk(cost, post_risk)
    cost_opt = tmp[1]; risk_min = tmp[2]
    return(cost_opt)
}) # end sapply
```

```{r}
length(cost_opts)
```


reproducing figure 2 in the lecture 02
```{r}
plot( sum_x, cost_opts, type = "l",    col = "blue", ylim = c(0, 1))
lines(sum_x, cost_opts,                col = "blue")
lines(sum_x, sum_x / n,               col = "chartreuse4")
lines(sum_x, rep(0.1, length(sum_x)), col = "red")

points(sum_x, cost_opts,                pch = 20, col = "blue")
points(sum_x, sum_x / n,               pch = 20, col = "chartreuse4")
points(sum_x, rep(0.1, length(sum_x)), pch = 20, col = "red")

legend("topleft", 
       lty = c(1,1,1), 
       col = c("blue", "chartreuse4", "red"), 
       cex = 0.8, 
       legend = c("Bayes Procedure (a = 0.05, b = 1)", "Sample Mean", "constant"))
```


```{r}
### data
sum_x = seq(0, 30, by = 1)
n = 30

### prior parameters:
as = c(0.05, 1, 2); bs = c(0.05, 1, 2)
shape_prior = expand.grid(as, bs)

### generate posterior risk
cost = seq(0, 1.0, by = 0.01)
s    = 100
set.seed(123)

mat_cost_opt = apply(shape_prior, 1, function(shape){
    cost_opt = sapply(sum_x, function(x){
        ### generate posterior risk
        post_risk = apply(as.matrix(cost), 1, posterior_risk, shape, x, n, s)
    
        ### get optimal cost
        tmp = cost_min_risk(cost, post_risk)
        cost_opt = tmp[1]; risk_min = tmp[2]
        return(cost_opt)
    }) # end sapply
    return(cost_opt)
}) # end apply
```

```{r}
shape_prior
```


```{r}
mat_cost_opt = t(mat_cost_opt)
```

```{r}
dim(mat_cost_opt)
```


```{r}
library(RColorBrewer)
```

```{r}
cols = brewer.pal(9, "Set1")
```

```{r}
tmp = apply(shape_prior, 1, function(shape){paste("Bayes:", "a =", shape[1], ";", "b =", shape[2])})
```


```{r}
plot( sum_x, sum_x / n, lwd = 2, lty = 2, type = "l", col = "black", ylim = c(0, 1), ylab = "Optimal Cost")
lines(sum_x, rep(0.1, length(sum_x)), lwd = 2, lty = 2, col = "grey50")

for (idx in 1:9){
    lines(sum_x, mat_cost_opt[idx,], col = cols[idx])
}

tmp = apply(shape_prior, 1, function(shape){paste("Bayes:", "a =", shape[1], ";", "b =", shape[2])})
legend("topleft", 
       lty = c(2, 2, rep(1, length(cols))), 
       col = c("black", "grey50", cols), 
       cex = 0.8, 
       legend = c("Sample Mean", "constant", tmp))
```


**The Bayes procedure always picks cost to be a little bigger than $\bar{x}$.**


-----

\color{black}  

## Task 4

```{r}
frequentist_risk = function(theta, loss_fun, find_optimal_cost, ...){
    ### generate different data x
    set.seed(123)
    sum_xs = rbinom(100, 30, theta)
    
    ### for different data, calculate optimal cost
    cost_opt = apply(as.matrix(sum_xs), 1, find_optimal_cost, ...)
  
    ### 
    loss = apply(as.matrix(cost_opt), 1, loss_fun, theta = theta)
    freq_risk = mean(loss)
    return(freq_risk)
}
```


```{r}
proc_bayes = function(sum_x, shape_prior, n, s = 100){
   ### generate posterior risk
    post_risk = apply(as.matrix(cost), 1, posterior_risk, shape_prior, sum_x, n, s)
    
    ### get optimal cost
    tmp = cost_min_risk(cost, post_risk)
    cost_opt = tmp[1]; risk_min = tmp[2]
    return(cost_opt)
} # end fun
```

```{r}
a = 0.05; b = 1

thetas = seq(0, 1, by=0.1)
risk = sapply(thetas, function(theta){
    frequentist_risk(theta, loss_fun, proc_bayes, n = 30, shape_prior = c(a, b))
}) # end sapply
```

```{r}
proc_mean = function(sum_x, n){
    return(sum_x / n)
}

risk2 = sapply(thetas, function(theta){
    frequentist_risk(theta, loss_fun, proc_mean, n = 30)
})
```


```{r}
proc_const = function(sum_x, n, const = 0.5){
    return(const)
}

risk3 = sapply(thetas, function(theta){
    frequentist_risk(theta, loss_fun, proc_const, n = 30, const = 0.1)
})
```

```{r}
plot(thetas, risk, type = "l", ylim = c(0,1), col = "grey50")
lines(thetas, R[1, ],  col = "green")
lines(thetas, risk2,   col = "grey50")
lines(thetas, R[2, ],  col = "blue")
lines(thetas, risk3,   col = "grey50")
lines(thetas, R[3, ],  col = "red")

```


For all $\theta$, the Bayes procedure has the lower frequentist risk than the sample mean. 
```{r, cache=TRUE}

thetas = seq(0, 1, by=0.1)

# frequentist risk for the 3 estimators given a theta
find_optimal_C = function(sum_x, a_prior, b_prior, n, s = 500){
  cost = seq(0, 1, by = 0.01)
  post_risk =  apply(as.matrix(cost), 1, posterior_risk, c(a_prior, b_prior), sum_x, n, s)
  cost[which.min(post_risk)]
}

frequentist_risk = function(theta){
    set.seed(123)
  sum_xs = rbinom(100, 30, theta)
  Bayes_optimal = apply(as.matrix(sum_xs), 1, find_optimal_C, a, b, n, s = 100)
  mean_c = sum_xs / 30
  
  loss1 = apply(as.matrix(Bayes_optimal), 1, loss_fun, theta = theta)
  loss2 = apply(as.matrix(mean_c), 1, loss_fun, theta = theta )
  risk1 = mean(loss1)
  risk2 = mean(loss2)
  risk3 = loss_fun(theta, 0.1)
  return(c(risk1, risk2, risk3))
}

# given a sequance a theta, compute frequentist risk for each theta
R = apply(as.matrix(thetas), 1, frequentist_risk)

# plot
plot(thetas, R[1,], col='blue', type = "l", 
     ylab = "frequentist risk", xlab = 'theta',ylim = c(0,1))
par(new = T)
plot(thetas, R[2,], type = 'l', col='green', 
     ylab = "", xlab = '', ylim = c(0,1))
par(new = T)
plot(thetas, R[3,], type = 'l',col = 'red',
     ylab = "", xlab = '', ylim = c(0,1))
legend("topright", lty = c(1,1,1), col = c("blue", "green", "red"),
       legend = c("Bayes", "Sample mean", "constant"))
```


\color{black}

## Task 5

